---
title: "Analysis: Urn Sampling (Combinatorics)"
author: "Mateusz Szczukiewicz"
date: "`r Sys.Date()`"
output: 
  pdf_document:   
    toc: true
    latex_engine: xelatex
  html_document: 
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 10, fig.height = 6)
library(ggplot2)
library(readr)
library(glue)
```

# 1. Problem Description

An urn contains 3 white balls (W) and 7 blue balls (B). We draw 3 balls without replacement. What is the probability of drawing at least two blue balls?

# 2. Analytical Solution (C++)

We use the combinatorics library, specifically the combinations function (based on the Symbol of Newton), to calculate the exact probability using the classical definition P(A)=|A|/|Ω|. We run the compiled analytics.cpp program.

```{r analytics}
analytics_exe <- file.path("..", "..", "build", "tasks", "combinatorics_urn_sampling", "combinatorics_urn_sampling_analytics")

analytics_result_str <- tail(system(analytics_exe, intern = TRUE), 1)
analytics_result_num <- as.numeric(analytics_result_str)

analytical_fraction <- "49/60"
```

The analytical probability P(A) is `r analytics_result_str` (exactly `r analytical_fraction`).

```cpp
// analytics.cpp snippet:
// long long omega = Stats::Combinatorics::combinations(10, 3);
// long long successes = Stats::Combinatorics::combinations(7, 2) * Stats::Combinatorics::combinations(3, 1) + 
//                       Stats::Combinatorics::combinations(7, 3) * Stats::Combinatorics::combinations(3, 0);
// double probability = static_cast<double>(successes) / omega;
// std::cout << probability << std::endl;
```

# 3. Simulation Verification (C++ & R)

We run the simulation.cpp program, which performs a Monte Carlo simulation of the urn drawing experiment `r {N_TRIALS <- 1e6; format(N_TRIALS, scientific = FALSE)}` times. The C++ program saves the convergence data (estimated probability after intervals of trials) to sim_data.csv.

```{r simulation}
simulation_exe <- file.path("..", "..", "build", "tasks", "combinatorics_urn_sampling", "combinatorics_urn_sampling_simulation")

system(simulation_exe)

sim_data <- read_csv("sim_data.csv")

final_sim_result <- tail(sim_data$P_A, 1)
total_trials <- tail(sim_data$Trial, 1)
```

Simulation result (after `r format(total_trials, scientific = FALSE)` trials): `r final_sim_result`

# 4. Convergence Visualization (R)

The following plot, generated in R using ggplot2, shows how the estimated probability P(A) from the Monte Carlo simulation (blue line) converges towards the exact analytical result (red dashed line) as the number of trials increases.

```{r convergence_plot}
ggplot(sim_data, aes(x = Trial, y = P_A)) +
    geom_line(color = "steelblue", alpha = 0.8) +
    geom_hline(aes(yintercept = analytics_result_num), color = "red", linetype = "dashed", size = 1) +
    scale_x_continuous(labels = scales::comma) +
    labs(
        title = "Monte Carlo Simulation Convergence (Urn Sampling)",
        subtitle = "Problem: P(>= 2 Blue | Draw 3 from 3W, 7B without replacement)",
        x = "Number of Trials",
        y = "Estimated P(A)",
        caption = glue("Red Dashed Line: Analytical Result ({analytical_fraction} ≈ {round(analytics_result_num, 4)})")
    ) +
    theme_minimal(base_size = 12) +
    theme(plot.caption = element_text(hjust = 0))
```

# 5. Conclusion

The analytical result for the probability of drawing at least two blue balls, calculated as `r round(analytics_result_num, 4)` (`r analytical_fraction`), was successfully verified by the Monte Carlo simulation. The simulation, after `r format(total_trials, scientific = FALSE)` trials, converged to a value of `r round(final_sim_result, 4)`, closely matching the theoretical prediction. The convergence plot visually confirms this agreement.
